Towards a Pandemic-Scale Ancestral Recombination Graph of SARS-CoV-2

### 2023-05-03 paper.tex
Yans' updates

\subsection{Backbone phylogeny}
For validation purposes, we compare the genetic
relationships in the sc2ts-inferred ARGs with phylogenetic relationships
inferred by the NextStrain project \citep{Hadfield2018-ef}. The NextStrain
trees consist of subsamples of SARS-CoV-2 genomes, which over the time period
of our ARGs are almost entirely restricted to non-recombinant lineages. To
allow for comparison, the ARGs were ``simplified'' \citep{Kelleher2018-xc} to
retain only those samples of our genealogy that match those present in a
specific NextStrain tree. These simplified ``backbone'' ARGs exhibit very
little recombination (6 and 5 trees across the genome for the wide and long
ARGs, respectively).

@@

\subsection{Tree inference from HMM sample clusters}
\label{sec:sample-cluster-tree-inference}
For a given daily batch of sample sequences that are to be attached to the
ARG, the LS HMM provides the most likely copying path and required mutations
for each of those sequences. We could directly attach these samples to
the ARG according to the copying paths and sets of mutations, but
such an approach would be highly unparsimonious because there are
often many samples that share the same copying path, with
complex patterns of shared mutations among those clusters of samples.
In this section we discuss the strategy used to infer likely tree
topologies within these clusters of samples sharing the same copying path.
In the majority of cases the copying path will consist of a single
node (i.e., no recombination), and so for simplicity we will assume in the
following discussion that all of the samples in a HMM-defined cluster
attach to a single node. The special treatment afforded to
newly-detected putative recombinants are discussed below
in Section XXX.

>>>>

\subsection{Backbone phylogeny}
For validation purposes, we compare the genetic
relationships in the sc2ts-inferred ARGs with phylogenetic relationships
inferred by the NextStrain project \citep{Hadfield2018-ef}. The NextStrain
trees consist of subsamples of SARS-CoV-2 genomes, which over the time period
of our ARGs are almost entirely restricted to non-recombinant lineages. To
enable this comparison, the ARGs were ``simplified'' \citep{Kelleher2018-xc} to
retain only those samples of our genealogy that match those present in a
specific NextStrain tree. These simplified ``backbone'' ARGs exhibit very
little recombination (6 and 5 trees across the genome for the wide and long
ARGs, respectively).

@@

\subsection{Tree inference from HMM daily sample clusters}
\label{sec:sample-cluster-tree-inference}
For a given daily batch of sample sequences that are to be attached to the
ARG, the LS HMM provides the most likely copying path and required mutations
for each of those sequences. We could directly attach these samples to
the ARG according to the copying paths and sets of mutations, but
such an approach would be highly unparsimonious because there are
often many samples that share the same copying path, with
complex patterns of shared mutations among those clusters of samples.
In this section we discuss the strategy used to infer likely tree
topologies within these clusters of samples sharing the same copying path.
In the majority of cases the copying path will consist of a single
node (i.e., no recombination), and so for simplicity we will assume in the
following discussion that all of the samples in a HMM-defined cluster
attach to a single node. The special treatment afforded to
newly-detected putative recombinants are discussed below
in Section XXX.
